Scikit-learn: machine learning without learning the machinery

### 2014-07-01 article.tex
restructure big data section

\paragraph{What about big data?}
%
``Big data'' is the latest trend of data analytics: applying machine
learning to large datasets can extract information so far unexploited.
These large datasets present new technical challenges, in
particular to be computationally tractable. It is important to
distinguish between the case of very large number of samples and the case of large
number of features. For a very large number of features, on-the-fly data
reductions, that exploit latent redundancy in the features, come in
handy. In scikit-learn, one could for instance use
random projections, that exploit the Johnson-Lindenstrauss lemma to give
good approximations of Euclidean distances in the dataset. For
a very large number of samples, some estimators expose
\emph{on-line} algorithms: they can be fit incrementally, with successive
calls to their {\tt partial\_fit} method, and converge in a small number
of passes on the data. To alleviate one of the main practical challenges of
big data, data-transfer latency, optimal data-fetching code can be used
in parallel with the calls to {\tt partial\_fit}.

Dealing with a large corpus of text faces both challenges: many samples,
and a very high dimensional space. In addition, the dimensionality
tends to increase with sample size, as the number of different words used
is greater in large corpuses. Beyond the mere size of the data matrix
created by a standard vectorization approach, as above, another challenge
is that the vectorization must be done in two passes over the documents,
as the size of the output vectors is not known beforehand.
Here, \emph{feature hashing} can be used
to tackle both of these challenges and produce a streaming vectorizer,
the {\tt HashingVectorizer} in scikit-learn.
The idea is to run words through a hash function, rather than a lookup table,
to map words to indices.
Collisions may occur, but the hashing algorithm \cite{weinberger2009}
is specially designed to mitigate their effect.
The dimensionality of the output vectors can be set by the user;
the bigger, the less likely collisions will occur.

>>>>

\paragraph{What about big data?}
%
\emph{Big data} is a major trend in data analytics: applying machine
learning to large datasets can extract information so far unexploited.
While scikit-learn is most efficient in the ``medium data'' range,
using shared memory multiprocessors rather than clusters,
it has facilities for dealing with datasets that don't fit
in any single machine.

Thinking in terms of the $(n \times p)$ data matrix,
we can distinguish between the case of large numbers of samples $n$,
and large numbers of features $p$.
For a large number of features, scikit-learn's options include
random projections, which reduce data dimensionality
while preserving most of the geometric structure.
For the ``large $n$'' case, some estimators expose a \texttt{partial\_fit}
method that does \emph{online learning}.
Instead of stashing all of the data in a single matrix,
the user can feed the estimator slices of the matrix,
and the algorithm finds a good approximation to its objective
after a few passes over the dataset.

% The citation is for "Heaps' law":
% vocabulary size as a function of collection size.
% Gael, feel free to remove it if you think it's too detailed.
With large collections of text, one faces both challenges:
the dimensionality $p$
tends to increase with sample size $n$, as the number of different words used
is greater in large collections \cite[88--89]{manning2009}.
Beyond the mere size of the data matrix
created by a standard vectorization approach, outlined earlier, another challenge
is that the vectorization must be done in two passes over the documents,
as the size of the output vectors is not known up front.
Here, \emph{feature hashing} can be used
to tackle both of these challenges and produce a streaming vectorizer,
the {\tt HashingVectorizer} in scikit-learn.
Words are run through a (stateless) hash function, rather than a lookup table,
to map words to indices.
Collisions may occur, but the hashing algorithm \cite{weinberger2009}
is specially designed to mitigate their effect.
The dimensionality of the output vectors can be set by the user;
the bigger, the less likely collisions will occur.
